Data Preparation

### 2019-01-11 Data Preparation.Rmd
debug student data

```{r import d_transcript, cache = TRUE}
d_transcript <- rbind(read_csv("Raw Grades/grades1.csv", col_types = cols()),
                      read_csv("Raw Grades/grades2.csv", col_types = cols())) %>%
  distinct
```

@@

```{r extracting overview}
# Setup
n_catalogue <- length(corpus_catalogues)
calendar_years <- c("2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019")
d_description <- tibble(Code            = character(0),
                        `Calendar Year` = character(0),
                        Overview        = character(0))

@@

```{r extracting description}
# Setup
symbol_to_line <- c("\r\n", "\n") %>%
  paste(collapse = "|")

@@

To put everything into tidy text format, we first create three tibbles `d_overview_tidy`, `d_description_tidy`, `d_manual` that respectively store the overviews, descriptions and text from manuals in the tidy text format, with one row per course-year-word (and course-word for `d_manual`).
```{r tidy d_description}
d_overview <- d_description %>%
  select(Code, `Calendar Year`, Overview) %>%
  unnest_tokens(output = word, input = Overview)
print(d_overview_tidy)

@@

# terms for which the stem differs from the original word
filter(dictionary, word != word_stem)

my_sw <- c("______________________________________________________________________________")

data("stop_words")  #imports common stopwords

my_stop_words <- data.frame (word = c(1:11, "students", )) #no stemming ("students")

manuals_tokens_preprocessed <- manuals_tokens %>%
  anti_join(stop_words)     %>%
  anti_join(my_stop_words) %>%
  anti_join(number_stopwords, by = c("word"="word"))

```{r stemming d_description}
d_overview <- d_overview %>%
  left_join(dictionary, by = "word") %>%
  rename(word_original = word,
         word = word_stem) %>%
  filter(!word %in% stop_words$word)

d_description <- d_description %>%
  left_join(dictionary, by = "word") %>%
  rename(word_original = word,
         word = word_stem) %>%
  filter(!word %in% stop_words$word)

rm(stems, stem_hunspell)
print(d_overview_stem) # See humanities - humanity
```
### Removing Stopwords

@@

```{r}
d_transcript <- d_transcript %>%
  filter(`Appraisal (Description)` == "Grade supervisor")
  select(`Student Number`, `Module (Abbrev.)`, `Module (Desc.)`,
         `Program (Description)`, `Academic Year`, `Academic Session`,
         `Number rep. attemp`, `Grade symbol`, `Booking Status (Desc.)`) %>%
  distinct %>%
  rename(`Student ID Number` = `Student Number`,
         `Course Code` = `Module (Abbrev.)`,
         `Course Title` = `Module (Desc.)`,
         `Study Program` = `Program (Description)`,
         `Civil Year` = `Academic Year`, # double check
         `Period` = `Academic Session`,
         `Number Repeated Attempt` = `Number rep. attemp`,
         `Grade` = `Grade symbol`,
         `Status` = `Booking Status (Desc.)`) # is there a better name for the variable?
```

Each faculty had a different code for their calendars, we want to standardise them so we can compare. We convert everything into UCM calendar codes (the correspondance between calendars was given to us by R. Vos) and save it under a variable `Period`:
```{r}
d_transcript <- d_transcript %>%
  mutate(Period = case_when(Period == 1 ~ "1 to 6"  ,
                            Period == 2 ~ "1 to 3"  ,
                            Period == 3 ~ "4 to 6"  ,
                            Period == 680 ~ "5"     , # exception
                            Period == 350 ~ "3 to 5", # exception
                            Period == 403 ~ "4 to 5", # exception
                            Period == 900 ~ "1 to 6", # exception
                            between(Period, 100, 199) ~ "1",
                            between(Period, 200, 299) ~ "2",
                            between(Period, 300, 399) ~ "3",
                            between(Period, 400, 499) ~ "4",
                            between(Period, 500, 599) ~ "5",
                            between(Period, 600, 660) ~ "6",
                            between(Period, 901, 999) ~ "1"))

###Extracting grades
We notice that there are several values under `Grade` for the same course. We are interested in keeping only the final grades, corresponding to the rows where `Appraisal Title` is marked as "Grade supervisor". 
So lets extract the grades and add the course concentration by keeping the first three letters of the `Code`:

@@

```{r}
d_student_concentration <- d_grade %>%
  group_by(ID, `Course concentration`) %>%
  summarise(count = n()) %>%
  arrange(ID,desc(count)) %>% #LONG
  spread(`Course concentration`,count, fill=0) %>% #WIDE
  transmute(student_Concentration = case_when(
    (HUM  >  2    &   SSC >  2   &  SCI <= 2) ~ "HUM/SSC",
    (HUM  >  2    &   SSC <= 2   &  SCI >  2) ~ "HUM/SCI",
    (HUM  >  2    &   SSC <= 2   &  SCI <= 2) ~ "HUM",
    (HUM  <= 2    &   SSC >  2   &  SCI >  2)  ~ "SSC/SCI",
    (HUM  <= 2    &   SSC >  2   &  SCI <= 2) ~ "SSC",
    (HUM  <= 2    &   SSC <= 2   &  SCI >  2)  ~ "SCI",
    (HUM  <  2    &   SSC <  2   &  SCI <  2) ~ "UNDECIDED",
    (HUM  >  2    &   SSC >  2   &  SCI >  2) ~ "CONFUSED", #normal cases ^
    (HUM > SCI    &   HUM > SSC)              ~ "HUM",
    (SSC > SCI    &   SSC > HUM)              ~ "SSC",
    (SCI > HUM    &   SCI > SSC)              ~ "HUM",
    TRUE ~ "oops"
  ))
```

Since we will need to check each student with an "oops" concentration to see what happened, lets separate this data for later:
```{r}
checkup <- filter(d_student_concentration, student_Concentration == "oops") 
```

@@

First, lets find out, which year was a student's first year and which year was a student's last year (we will also create a variable `Subs_year` that we will substract later to transform from Calendar Year to UCM Year, and a second variable `LengthUCM` that shows us how many years each student stayed at the college):
```{r}

@@

@@

```{r}
var_years <- as.character(2007:2017)

@@

```{r}
d_students <- d_grade %>%
  select(ID, Code, Year, Period, Grade, Attempt,`Booking Status Description`, `Course concentration`) %>%
  arrange(ID, Year, Period)%>%
  left_join(d_student_concentration, by="ID") %>%
  mutate(Grade = as.numeric( sub(",",".", Grade))) %>%
  replace_na(list(Grade=-1))

@@

For this we keep only the capstones. We expect some people to have repeated capstones:
```{r}
capstone_count <- d_students %>%
  filter(Code=="CAP3000") %>%
  group_by(ID)%>%
  summarise(Count=n())

@@

```{r}
passed_capstone_count <- d_students %>%
  filter(Code=="CAP3000", `Booking Status Description`=="Completed with Success") %>%
  group_by(ID) %>%
  summarise(Count=n(), Max = max(Grade), Min = min(Grade), Attempts= max(Attempt))

@@

We can check the grades for people who had repeated passed capstones (lets just bring them to the top of the list):
```{r}
passed_capstone_count %>%
  arrange(desc(Count), desc(Min), Attempts)
```
There are 1712 people who did capstone. We see that most students who have two passed Capstones, have in reality only 1, since the other capstone had a failing grade and the passing grade was on the second attempt (count starts at 0) was indeed failed. However there are two students who indeed have two different passed Capstones: one is student 282995 who has two different passing grades(because of a change in grading system at the collegue during the year 2008-2009), the other is student 544450 who has the same grade passed at two different moment in time.

We can check if there are any students with passing capstone who do not have a grade of at least 5.5. We do this by looking at the lowest Max values:
```{r}
passed_capstone_count %>%
  arrange(Max)
```
We see the lowest grade for a passing capstone is 5.6, so this is fine.

Now lets just keep the highest value for each student and correct for student 282995:
```{r}
passed_capstone_grades <- passed_capstone_count %>%
  mutate(`Booking Status Description`= "Completed with Success", Code="CAP3000", Grade= case_when(ID== 282995~7.7,
                                                                                  T~Max)) %>%
  select( ID, Code, Grade,`Booking Status Description`)
passed_capstone_grades
```

@@

First we get the years for which there is a corresponding passing capstone grade, and check for repetition:
```{r}

@@

Lets create a dataframe where we include the capstone year. Then, if a course is given later than the capstone year, we want to remove it. Therefore, lets substitute na values with the year 9999 (or will ucm be open in the year 9999?).
```{r}
d_transcripts <- d_students %>%
  group_by(ID) %>%
  left_join(capstone_finish_year) %>%
  ungroup() %>%
  replace_na(list(`Capstone Year`=9999)) %>%
  mutate(`After Graduation`=`Capstone Year`-`Year`)

d_transcripts
```
Now we will remove all courses with negative `After Graduation`:
```{r}
master_courses <- d_transcripts%>%
  filter(`After Graduation`<0)

@@

```{r}
booked <- filter(d_transcripts, d_transcripts$`Booking Status Description`=="Booked")%>%
  mutate(`Booking Status Description`= case_when(Grade<5.5~"Completed Unsuccessfully",
                                                 Grade>5.5~"Completed with Success"))

@@

# Save Data
```{r}
save(lists, d_course, d_AoD, d_assessment, d_transcript,
     file = "data_pillar_1.RDATA")
save(lists, d_course, d_AoD, d_assessment, d_transcript,
     d_overview_stem, d_description_stem,
     file = "data_pillar_2.RDATA")
```

>>>>

```{r import d_transcript, cache = TRUE}
d_transcript <- rbind(read_csv("Raw Grades/grades1.csv", col_types = cols()),
                      read_csv("Raw Grades/grades2.csv", col_types = cols()))
```

@@

```{r extracting overview, cache = TRUE}
# Setup
n_catalogue <- length(corpus_catalogues)
calendar_years <- c("2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019")
d_description <- tibble(Code            = character(0),
                        `Calendar Year` = character(0),
                        Overview        = character(0))

@@

```{r extracting description, cache = TRUE}
# Setup
symbol_to_line <- c("\r\n", "\n") %>%
  paste(collapse = "|")

@@

To put everything into tidy text format, we first create three tibbles `d_overview_tidy`, `d_description_tidy`, `d_manual` that respectively store the overviews, descriptions and text from manuals in the tidy text format, with one row per course-year-word (and course-word for `d_manual`).
```{r tidy d_description}
d_overview <- d_description %>%
  select(Code, `Calendar Year`, Overview) %>%
  unnest_tokens(output = word, input = Overview)

@@

# terms for which the stem differs from the original word
filter(dictionary, word != word_stem)
```

```{r stemming d_description}
stem_with_dictionary <- function(data) data %>%
                                         left_join(dictionary, by = "word") %>%
                                         rename(word_original = word,
                                                word = word_stem)

d_description <- stem_with_dictionary(d_description)
d_overview <- stem_with_dictionary(d_overview)
d_manual <- stem_with_dictionary(d_manual)

rm(stems, stem_hunspell)
print(d_overview_stem) # See humanities - humanity
```
### Removing Stopwords
```{r}

# Own stop words
sw_own <- c("______________________________________________________________________________")

remove_sw <- function(data) data %>%
  filter(! word %in% c(stop_words$word, sw_own))

d_description <- remove_sw(d_description)
d_overview <- remove_sw(d_overview)
d_manual <- remove_sw(d_manual)
```

@@

```{r}
d_transcript <- d_transcript %>%
  filter(`Appraisal (Description)` == "Grade supervisor") %>%
  select(`Student Number`, `Module (Abbrev.)`, `Module (Desc.)`,
         `Program (Description)`, `Academic Year`, `Academic Session`,
         `Number rep. attemp`, `Grade symbol`, `Booking Status (Desc.)`) %>%
  rename(`Student ID Number` = `Student Number`,
         `Course Code` = `Module (Abbrev.)`,
         `Course Title` = `Module (Desc.)`,
         `Study Program` = `Program (Description)`,
         `Civil Year` = `Academic Year`, # double check
         `Period` = `Academic Session`,
         `Number Repeated Attempt` = `Number rep. attemp`,
         Grade = `Grade symbol`,
         `Status` = `Booking Status (Desc.)`) %>% # is there a better name for the variable?
  distinct
```

Each faculty had a different code for their calendars, we want to standardise them so we can compare. We convert everything into UCM calendar codes and save it under a variable `Period`:
```{r}
d_transcript <- d_transcript %>%
  mutate(Period = case_when(Period == 1 ~ "1 to 6"  ,
                            Period == 2 ~ "1 to 3"  ,
                            Period == 3 ~ "4 to 6"  ,
                            Period == 680 ~ "5"     , # exception
                            Period == 350 ~ "3 to 5", # exception
                            Period == 403 ~ "4 to 5", # exception
                            Period == 900 ~ "1 to 6", # exception
                            between(Period, 100, 199) ~ "1",
                            between(Period, 200, 299) ~ "2",
                            between(Period, 300, 399) ~ "3",
                            between(Period, 400, 499) ~ "4",
                            between(Period, 500, 599) ~ "5",
                            between(Period, 600, 660) ~ "6",
                            between(Period, 901, 999) ~ "1"))

```
###Dropping useless variables:
Now that our `Period` variable is clean, we can exclude the variables that we will not use, and we will delete newly generated duplicate rows:

```
###Extracting grades
We notice that there are several values under `Grade` for the same course. We are interested in keeping only the final grades, corresponding to the rows where `Appraisal Title` is marked as "Grade supervisor". 
So lets extract the grades and add the course concentration by keeping the first three letters of the `Code`:

@@

```{r, eval = FALSE}
d_concentration <- d_course %>%
  select(Code, Concentration, `Concentration (additional)`) %>%
  gather(key = X, value = Concentration, Concentration, `Concentration (additional)`, na.rm = TRUE) %>%
  left_join(d_transcript, by = c("Code" = "Course Code")) %>%
  select(Concentration, `Student ID Number`) %>%
  count(Concentration, `Student ID Number`) %>%
  spread(key = Concentration, value = n, fill = 0) %>%
  mutate(n_course = HUM + SCI + SSC,
         `Student Concentration` = case_when(
           n_course < 20 ~ "not enough courses",
           HUM > 5 & SSC > 5 & SCI > 5 ~ "confused",
           HUM > 5 & SSC > 5           ~ "HUM/SCI",
           HUM > 5                     ~ "HUM",
                     SSC > 5 & SCI > 5 ~ "SSC/SCI",
                     SSC > 5           ~ "SSC",
                               SCI > 5 ~ "SCI"
  ))

  transmute(student_Concentration = case_when(
    (HUM  >  2    &   SSC >  2   &  SCI <= 2) ~ "HUM/SSC",
    (HUM  >  2    &   SSC <= 2   &  SCI >  2) ~ "HUM/SCI",
    (HUM  >  2    &   SSC <= 2   &  SCI <= 2) ~ "HUM",
    (HUM  <= 2    &   SSC >  2   &  SCI >  2)  ~ "SSC/SCI",
    (HUM  <= 2    &   SSC >  2   &  SCI <= 2) ~ "SSC",
    (HUM  <= 2    &   SSC <= 2   &  SCI >  2)  ~ "SCI",
    (HUM  <  2    &   SSC <  2   &  SCI <  2) ~ "UNDECIDED",
    (HUM  >  2    &   SSC >  2   &  SCI >  2) ~ "CONFUSED", #normal cases ^
    (HUM > SCI    &   HUM > SSC)              ~ "HUM",
    (SSC > SCI    &   SSC > HUM)              ~ "SSC",
    (SCI > HUM    &   SCI > SSC)              ~ "HUM",
    TRUE ~ "oops"
  ))
```

Since we will need to check each student with an "oops" concentration to see what happened, lets separate this data for later:
```{r, eval = FALSE}
checkup <- filter(d_student_concentration, student_Concentration == "oops") 
```

@@

First, lets find out, which year was a student's first year and which year was a student's last year (we will also create a variable `Subs_year` that we will substract later to transform from Calendar Year to UCM Year, and a second variable `LengthUCM` that shows us how many years each student stayed at the college):
```{r, eval = FALSE}

@@

```{r, eval = FALSE}

@@

```{r, eval = FALSE}
var_years <- as.character(2007:2017)

@@

```{r, eval = FALSE}
d_students <- d_grade %>%
  select(ID, Code, Year, Period, Grade, Attempt,`Booking Status Description`, `Course concentration`) %>%
  arrange(ID, Year, Period)%>%
  left_join(d_student_concentration, by="ID") %>%
  mutate(Grade = as.numeric( sub(",",".", Grade))) %>%
  replace_na(list(Grade=-1))

@@

For this we keep only the capstones. We expect some people to have repeated capstones:
```{r, eval = FALSE}
capstone_count <- d_students %>%
  filter(Code=="CAP3000") %>%
  group_by(ID)%>%
  summarise(Count=n())

@@

```{r, eval = FALSE}
passed_capstone_count <- d_students %>%
  filter(Code=="CAP3000", `Booking Status Description`=="Completed with Success") %>%
  group_by(ID) %>%
  summarise(Count=n(), Max = max(Grade), Min = min(Grade), Attempts= max(Attempt))

@@

We can check the grades for people who had repeated passed capstones (lets just bring them to the top of the list):
```{r, eval = FALSE}
passed_capstone_count %>%
  arrange(desc(Count), desc(Min), Attempts)
```
There are 1712 people who did capstone. We see that most students who have two passed Capstones, have in reality only 1, since the other capstone had a failing grade and the passing grade was on the second attempt (count starts at 0) was indeed failed. However there are two students who indeed have two different passed Capstones: one is student 282995 who has two different passing grades(because of a change in grading system at the collegue during the year 2008-2009), the other is student 544450 who has the same grade passed at two different moment in time.

We can check if there are any students with passing capstone who do not have a grade of at least 5.5. We do this by looking at the lowest Max values:
```{r, eval = FALSE}
passed_capstone_count %>%
  arrange(Max)
```
We see the lowest grade for a passing capstone is 5.6, so this is fine.

Now lets just keep the highest value for each student and correct for student 282995:
```{r, eval = FALSE}
passed_capstone_grades <- passed_capstone_count %>%
  mutate(`Booking Status Description`= "Completed with Success", Code="CAP3000", Grade= case_when(ID== 282995~7.7,
                                                                                  T~Max)) %>%
  select( ID, Code, Grade,`Booking Status Description`)
passed_capstone_grades
```

@@

First we get the years for which there is a corresponding passing capstone grade, and check for repetition:
```{r, eval = FALSE}

@@

Lets create a dataframe where we include the capstone year. Then, if a course is given later than the capstone year, we want to remove it. Therefore, lets substitute na values with the year 9999 (or will ucm be open in the year 9999?).
```{r, eval = FALSE}
d_transcripts <- d_students %>%
  group_by(ID) %>%
  left_join(capstone_finish_year) %>%
  ungroup() %>%
  replace_na(list(`Capstone Year`=9999)) %>%
  mutate(`After Graduation`=`Capstone Year`-`Year`)

d_transcripts
```
Now we will remove all courses with negative `After Graduation`:
```{r, eval = FALSE}
master_courses <- d_transcripts%>%
  filter(`After Graduation`<0)

@@

```{r, eval = FALSE}
booked <- filter(d_transcripts, d_transcripts$`Booking Status Description`=="Booked")%>%
  mutate(`Booking Status Description`= case_when(Grade<5.5~"Completed Unsuccessfully",
                                                 Grade>5.5~"Completed with Success"))

@@

# Save Data

```{r}
save(lists, d_course, d_AoD, d_assessment, d_transcript,
     file = "data_pillar_1.RDATA")
save(lists, d_course, d_AoD, d_assessment, d_transcript,
     d_overview, d_description, d_manual,
     file = "data_pillar_2.RDATA")
```
